---
title: "Wine Imports to U.S."
author: "Henry Johnson"
date: "3/1/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# wd <- "/Users/henryjohnson/Documents/school/thesis/the-effects-of-tariffs-on-wine-prices"

# setwd(wd)
library(foreign)
library(ggplot2)
library(estimatr)
library(car)
library(stats)
library(corrplot)
library(readxl)
library(stringr)
library(dplyr)
library(forecast)
library(panelvar)
library(tseries)
library(plm)
library(utils)
```

## Import data

```{r import-data}
test_for_r <- read_excel("~/Documents/school/Thesis/the-effect-of-tariffs-on-wine-prices/data/test_for_r.xlsx", 
     col_types = c("skip", "text", "date", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "skip", "numeric", "numeric", "numeric", 
         "numeric", "numeric", "numeric", 
         "numeric", "numeric"))
```

## Rename Columns

```{r rename-columns, echo=FALSE}
df <- test_for_r
names(df)[3] <- 'cif'
names(df)[4] <- 'duties'
names(df)[5] <- 'duty_val'
names(df)[6] <- 'duty_paid'
names(df)[7] <- 'quantity_imp'
names(df)[8] <- 'value_imp'
names(df)[9] <- 'ppi'
```


## Adjust for Inflation
```{r adjust-inflation}
df$cif_adj <- df$cif / df$ppi * 100
df$duties_adj <- df$duties / df$ppi * 100
df$duty_val_adj <- df$duty_val / df$ppi * 100
df$duty_paid_adj <- df$duty_paid / df$ppi * 100
df$value_imp_adj <- df$value_imp / df$ppi * 100
```


```{r create-synthetic-variables}
df$duty_rate <- ifelse(df$duty_val==0, 0, df$duties / df$duty_val)
df$imp_pc <- df$quantity_imp / df$population
df$bottled_pc <- df$bottled / df$population
df$cif_rate <- ifelse(df$quantity_imp==0, 0, df$cif_adj / df$quantity_imp)

# create proportion of imports column
df1 = subset(df, select=c("Country", "month", "quantity_imp"))
df1 <- df1[(df1$Country=="Victims"),]
df1 <- df1 %>% distinct()
df2 = subset(df, select=c("Country", "month", "quantity_imp"))
df2 <- df2[(df2$Country=="World"),]
df2 <- df2 %>% distinct()
df3 = subset(df, select=c("month"))
df3 <- df3 %>% distinct()
df3$prop_imp <- df1$quantity_imp / (df2$quantity_imp - df1$quantity_imp)
# merge proportions back in
df <- left_join(df, df3, by="month")

rm(df1)
rm(df2)
rm(df3)
```

## Stationary Corrections

First, I'll test for stationarity using ADF test and then diff the data to increase stationarity.

```{r stationarity}
# duty rate isn't stationary by itself for a couple of countries
for (c in unique(df$Country)) {
  test_data <- df[(df$Country==c),]
  test_data = subset(test_data, select=c("month", "duty_rate"))
  test_data <- test_data %>% distinct()
  print(c)
  print(adf.test(na.omit(test_data$duty_rate)))
}
# duty rate isn't stationary by itself for a couple of countries. Taking the first diff works.
df <- transform(df, duty_r_d1 = ave(duty_rate, Country, FUN = function(x) c(NA, diff(x))))
for (c in unique(df$Country)) {
  test_data <- df[(df$Country==c),]
  test_data = subset(test_data, select=c("month", "duty_r_d1"))
  test_data <- test_data %>% distinct()
  print(c)
  print(adf.test(na.omit(test_data$duty_r_d1)))
}

# cif rate isn't stationary by itself for a couple of countries
for (c in unique(df$Country)) {
  test_data <- df[(df$Country==c),]
  test_data = subset(test_data, select=c("month", "cif_rate"))
  test_data <- test_data %>% distinct()
  print(c)
  print(adf.test(na.omit(test_data$cif_rate)))
}
# taking the first diff works
df <- transform(df, cif_r_d1 = ave(cif_rate, Country, FUN = function(x) c(NA, diff(x))))
for (c in unique(df$Country)) {
  test_data <- df[(df$Country==c),]
  test_data = subset(test_data, select=c("month", "cif_r_d1"))
  test_data <- test_data %>% distinct()
  print(c)
  print(adf.test(na.omit(test_data$cif_r_d1)))
}

# values for bottled wine are stationary
test_data = subset(df, select=c("month", "bottled", "bottled_pc"))
test_data <- test_data %>% distinct()
adf.test(na.omit(test_data$bottled))
adf.test(na.omit(test_data$bottled_pc))

# price isn't stationary
test_data = subset(df, select=c("month", "price_adj"))
test_data <- test_data %>% distinct()
adf.test(na.omit(test_data$price_adj))
# first diff fixes that
price_diff <- diff(test_data$price_adj, lag=1, differences=1)
adf.test(na.omit(price_diff))
# merge it back to main df
price_diff <- c(NA, price_diff)
test_data$price_d1 <- price_diff
test_data <- subset(test_data, select=c("month", "price_d1"))
df <- left_join(df, test_data, by="month")

# export quantity isn't stationary by itself
test_data = subset(df, select=c("month", "quantity_exports"))
test_data <- test_data %>% distinct()
adf.test(na.omit(test_data$quantity_exports))
# first diff fixes that
exp_q_diff <- diff(test_data$quantity_exports, lag=1, differences=1)
adf.test(na.omit(exp_q_diff))
exp_q_diff <- c(NA, exp_q_diff)
test_data$exp_q_d1 <- exp_q_diff
test_data <- subset(test_data, select=c("month", "exp_q_d1"))
df <- left_join(df, test_data, by="month")

# proportion of imports is stationary by itself
test_data = subset(df, select=c("month", "prop_imp"))
test_data <- test_data %>% distinct()
adf.test(na.omit(test_data$prop_imp))

# quantity of imports isn't stationary since many countries aren't stationary
for (c in unique(df$Country)) {
  test_data <- df[(df$Country==c),]
  test_data = subset(test_data, select=c("month", "quantity_imp"))
  test_data <- test_data %>% distinct()
  print(c)
  print(adf.test(na.omit(test_data$quantity_imp)))
}
# taking the first diff works
df <- transform(df, imp_q_d1 = ave(quantity_imp, Country, FUN = function(x) c(NA, diff(x))))
for (c in unique(df$Country)) {
  test_data <- df[(df$Country==c),]
  test_data = subset(test_data, select=c("month", "imp_q_d1"))
  test_data <- test_data %>% distinct()
  print(c)
  print(adf.test(na.omit(test_data$imp_q_d1)))
}
```

## Correct for seasonality

```{r seasonality-imp-q-d1}
# Column: imp_q_d1
i=0
for (c in unique(df$Country)) {
  if (i == 0) {
    temp_df <- df[df$Country == c, ]

    df_diff_seas_quant_imp <- temp_df$imp_q_d1 %>%
      mstl()
    
    temp_ts <- ts(df_diff_seas_quant_imp, start=c(2000, 1), end=c(2021, 12), frequency=12)
    seas_adj <- seasadj(decompose(temp_ts, "multiplicative"))
    temp_df$imp_q_sa <- seas_adj[,1]
    temp_df = subset(temp_df, select=c("month", "Country", "imp_q_sa"))
    seas_adj_df <- left_join(df, temp_df, by = c("month", "Country"))
    seas_adj_df = seas_adj_df[, !(names(seas_adj_df) %in% c("imp_q_sa_2"))]
  }
  else {
    temp_df <- df[df$Country == c, ]
    
    df_diff_seas_quant_imp <- temp_df$imp_q_d1 %>%
      mstl()

    temp_ts <- ts(df_diff_seas_quant_imp, start=c(2000, 1), end=c(2021, 12), frequency=12)
    seas_adj <- seasadj(decompose(temp_ts, "multiplicative"))
    temp_df$imp_q_sa_2 <- seas_adj[,1]
    temp_df = subset(temp_df, select=c("month", "Country", "imp_q_sa_2"))
    seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month", "Country"))

    seas_adj_df$imp_q_sa <- coalesce(seas_adj_df$imp_q_sa, seas_adj_df$imp_q_sa_2)
    seas_adj_df = seas_adj_df[, !(names(seas_adj_df) %in% c("imp_q_sa_2"))]
  }
  i = i+1
}
```

```{r seasonality-cif-r-d1}
# Column: cif_r_d1
i=0
for (c in unique(df$Country)) {
  if (i == 0) {
    temp_df <- seas_adj_df[seas_adj_df$Country == c, ]

    df_diff_seas_quant_imp <- temp_df$cif_r_d1 %>%
      mstl()
    
    temp_ts <- ts(df_diff_seas_quant_imp, start=c(2000, 1), end=c(2021, 12), frequency=12)
    seas_adj <- seasadj(decompose(temp_ts, "multiplicative"))
    temp_df$cif_r_sa <- seas_adj[,1]
    temp_df = subset(temp_df, select=c("month", "Country", "cif_r_sa"))
    seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month", "Country"))
    seas_adj_df = seas_adj_df[, !(names(seas_adj_df) %in% c("cif_r_sa_2"))]
  }
  else {
    temp_df <- seas_adj_df[seas_adj_df$Country == c, ]
    
    df_diff_seas_quant_imp <- temp_df$cif_r_d1 %>%
      mstl()

    temp_ts <- ts(df_diff_seas_quant_imp, start=c(2000, 1), end=c(2021, 12), frequency=12)
    seas_adj <- seasadj(decompose(temp_ts, "multiplicative"))
    temp_df$cif_r_sa_2 <- seas_adj[,1]
    temp_df = subset(temp_df, select=c("month", "Country", "cif_r_sa_2"))
    seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month", "Country"))

    seas_adj_df$cif_r_sa <- coalesce(seas_adj_df$cif_r_sa, seas_adj_df$cif_r_sa_2)
    seas_adj_df = seas_adj_df[, !(names(seas_adj_df) %in% c("cif_r_sa_2"))]
  }
  i = i+1
}
```

```{r seasonality-duty-r-d1}
# Column: duty_r_d1
i=0
for (c in unique(df$Country)) {
  if (i == 0) {
    temp_df <- seas_adj_df[seas_adj_df$Country == c, ]

    df_diff_seas_quant_imp <- temp_df$duty_r_d1 %>%
      mstl()
    
    temp_ts <- ts(df_diff_seas_quant_imp, start=c(2000, 1), end=c(2021, 12), frequency=12)
    seas_adj <- seasadj(decompose(temp_ts, "additive"))
    temp_df$duty_r_sa <- seas_adj[,1]
    temp_df = subset(temp_df, select=c("month", "Country", "duty_r_sa"))
    seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month", "Country"))
    seas_adj_df = seas_adj_df[, !(names(seas_adj_df) %in% c("duty_r_sa_2"))]
  }
  else {
    temp_df <- df[df$Country == c, ]
    
    df_diff_seas_quant_imp <- temp_df$duty_r_d1 %>%
      mstl()

    temp_ts <- ts(df_diff_seas_quant_imp, start=c(2000, 1), end=c(2021, 12), frequency=12)
    seas_adj <- seasadj(decompose(temp_ts, "additive"))
    temp_df$duty_r_sa_2 <- seas_adj[,1]
    temp_df = subset(temp_df, select=c("month", "Country", "duty_r_sa_2"))
    seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month", "Country"))

    seas_adj_df$duty_r_sa <- coalesce(seas_adj_df$duty_r_sa, seas_adj_df$duty_r_sa_2)
    seas_adj_df = seas_adj_df[, !(names(seas_adj_df) %in% c("duty_r_sa_2"))]
  }
  i = i+1
}
```

```{r seasonality-exp-q-d1}
# Column: exp_q_d1
temp_df = subset(seas_adj_df, select=c("month", "exp_q_d1"))
temp_df <- temp_df %>% distinct()

df_diff_seas_exp_q_d1 <- temp_df$exp_q_d1 %>%
  mstl()

temp_ts <- ts(df_diff_seas_exp_q_d1, start=c(2000, 1), end=c(2021, 12), frequency=12)
temp_df$exp_q_sa <- seasadj(decompose(temp_ts, "multiplicative"))[,1]
temp_df = subset(temp_df, select=c("month", "exp_q_sa"))

seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month"))
```

```{r seasonality-price-d1}
# Column: price_d1
temp_df = subset(seas_adj_df, select=c("month", "price_d1"))
temp_df <- temp_df %>% distinct()

df_diff_seas_price_adj_d1 <- temp_df$price_d1 %>%
  mstl()

temp_ts <- ts(df_diff_seas_price_adj_d1, start=c(2000, 1), end=c(2021, 12), frequency=12)
temp_df$price_sa <- seasadj(decompose(temp_ts, "multiplicative"))[,1]
# temp_df$price_sa <- seasadj(df_diff_seas_price_adj_d1)
temp_df = subset(temp_df, select=c("month", "price_sa"))

seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month"))
```

```{r seasonality-bottled-pc}
# Column: bottled_d1
temp_df = subset(seas_adj_df, select=c("month", "bottled_pc"))
temp_df <- temp_df %>% distinct()

df_diff_bottled_pc_d1 <- temp_df$bottled_pc %>%
  mstl()

temp_ts <- ts(df_diff_bottled_pc_d1, start=c(2000, 1), end=c(2021, 12), frequency=12)
temp_df$bottled_pc_sa <- seasadj(decompose(temp_ts, "multiplicative"))[,1]
# temp_df$bottled_pc_sa <- seasadj(df_diff_bottled_pc_d1)
temp_df = subset(temp_df, select=c("month", "bottled_pc_sa"))

seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month"))
```

```{r seasonality-prop-imp}
# Column: prop_imp
temp_df = subset(seas_adj_df, select=c("month", "prop_imp"))
temp_df <- temp_df %>% distinct()

df_diff_prop_imp_d1 <- temp_df$prop_imp %>%
  mstl()

temp_ts <- ts(df_diff_prop_imp_d1, start=c(2000, 1), end=c(2021, 12), frequency=12)
temp_df$prop_imp_sa <- seasadj(decompose(temp_ts, "multiplicative"))[,1]
# temp_df$prop_imp_sa <- seasadj(df_diff_prop_imp_d1)
temp_df = subset(temp_df, select=c("month", "prop_imp_sa"))

seas_adj_df <- left_join(seas_adj_df, temp_df, by = c("month"))
seas_adj_df <- as.data.frame(seas_adj_df)

```

```{r cleanup-columns}
# seas_adj_df <- seas_adj_df[ , -which(names(seas_adj_df) %in% c("bottled_pc_sa.y","price_sa.y", "exp_q_sa.y"))]
# 
# seas_adj_df <- seas_adj_df %>%
#   rename(
#     bottled_pc_sa = bottled_pc_sa.x,
#     price_sa = price_sa.x,
#     exp_q_sa = exp_q_sa.x
#   )
# 
# seas_adj_df$price_sa
```

# Now let's do a panel VAR

```{r model-var}
model_df = subset(seas_adj_df, select=c("Country", "month", "price_d1", "bottled_pc_sa", "imp_q_sa", "duty_rate", "exp_q_sa", "frspger_25", "prop_imp_sa", "cif_r_sa"))

time_df <- model_df
time_df <- subset(time_df, select=c("month"))
time_df <- time_df %>% distinct()
seq <- seq(1, 261, by=1)
time_df$time_seq <- seq

model_df2 <- na.omit(model_df)

colnames(model_df2)
model_df2$Country = as.factor(model_df2$Country)

seas_adj_df$id <- as.integer(factor(seas_adj_df$Country, 
                  levels=unique(seas_adj_df$Country)))
model_df2$tariff <- as.logical(model_df2$frspger_25)
model_df2$time <- as.integer(model_df2$month)

model_df2$price_sa = as.numeric(model_df2$price_sa)
model_df2$bottled_pc_sa = as.numeric(model_df2$bottled_pc_sa)
model_df2$imp_q_sa = as.numeric(model_df2$imp_q_sa)
model_df2$exp_q_sa = as.numeric(model_df2$exp_q_sa)
model_df2$cif_r_sa = as.numeric(model_df2$cif_r_sa)

model_df3 <- model_df2[!(model_df2$Country=="World" | model_df2$Country=="Victims"),]

model_df2 <- left_join(model_df2, time_df, by="month")

# class(model_df2)
# names(model_df2)
model_df2 <- as.data.frame(model_df2)

```

Let's check to see if we might any multicollinearity issues by doing a corr plot
```{r corr-plots}

for (c in unique(model_df2$Country)) {
  temp_df <- model_df2[model_df2$Country == c, ]
  corr_plot_df <- subset(temp_df, select=c("price_d1", "bottled_pc_sa", "imp_q_sa", "duty_rate", "exp_q_sa", "cif_r_sa", "prop_imp_sa"))
  M <- cor(corr_plot_df)
  corrplot(M)
}

temp_df <- df[df$Country == "Victims", ]
corr_plot_df <- subset(temp_df, select=c("price_d1", "bottled_pc_sa", "imp_q_sa", "duty_r_d1", "exp_q_sa", "cif_r_sa", "prop_imp_sa"))
M <- cor(corr_plot_df)
corrplot(M)
```

There doesn't seem to be any correlation between duty rates and prices here. WTAF.

```{r fit-model}

# The dataframe becomes singular if we include all of the panels
model_df3 <- seas_adj_df[(model_df2$Country=="Victims" | model_df2$Country=="Argentina" |  model_df2$Country=="Chile" | model_df2$Country=="Italy" ),]

model_df3 <- as.data.frame(model_df3)

var_mod <-
  pvarfeols(
    dependent_vars = c("price_sa", "duty_r_sa", "exp_q_sa", "prop_imp_sa", "bottled_pc_sa"),
    lags = 1,
    exog_vars = c("cif_r_sa"),
    transformation = "demean",
    data = na.omit(model_df3),
    panel_identifier= c("Country", "month"))
summary(var_mod)


# fails due to singularity issue
model_df4 <- seas_adj_df[!(model_df2$Country=="Victims" | model_df2$Country=="World"),]

var_mod <-
  pvarfeols(
    dependent_vars = c("price_sa", "duty_r_sa", "exp_q_sa", "prop_imp_sa", "bottled_pc_sa"),
    lags = 1,
    exog_vars = c("cif_r_sa"),
    transformation = "demean",
    data = na.omit(model_df4),
    panel_identifier= c("Country", "month"))
summary(var_mod)

```


```{r irf}
library("MTS")
varx_irf <- VARXirf(var_mod, lag=12, orth=TRUE)

girf(var_mod, n.ahead=3, ma_approx_steps=3)
```


